---
title: "**Assignment 01 - Group 3**"
author: "Daniel + Megan + Mike"
date: "26 October 2021"
output: 
  rmdformats::downcute
---
<style>
body {
text-align: left}
</style>

<style>
body {
font-family: arial}
</style>

# research question
How well does a census tract’s percentage of single-family homes and location relative to metropolitan center predict the average home value and tenure period in homes for the Commonwealth of Massachusetts?

# prior research
While the correlation between diversity of housing types and racial and economic diversity in	neighborhoods have previously been established, this study investigated a potential 	relationship between housing typology and foreclosures. In fact, they found a positive (but varied) correlation between type and stability—neighborhoods with a range of housing stock had lower foreclosure and sales rates (Chakraborty et al 2018). Additionally, hedonic modelling of home values have been shown to respond to distance to city center. Longer distance from the metropolitan center correlates with lower property values in the magnitude of 1.82%. However, this relationship is highly income dependent, where devaluation associated with distance is higher for low-income individuals than for high-income individuals (Kestens et al. 2006).

- Chakraborty, Arnab and McMillan, Andrew, “Is Housing Diversity Good for Community Stability?: Evidence from the Housing Crisis,” *Journal of Planning Education and Research*, (2018), https://journals.sagepub.com/doi/10.1177/0739456X18810787
- Kestens, Yan,  Theriault, Marius, and Des Rosiers  Francois "Heterogeneity in hedonic modelling of house
prices: looking at buyers’ household profiles," *Journal of Geographical Systems*, (2006), https://link.springer.com/article/10.1007/s10109-005-0011-8

# data
The sample population for this study is the full set of all census tracts in the Commonwealth of Massachusetts. 

The analysis will include the following variables:

CONTINUOUS

- Percentage of Single, Detached Housing Units (United States Census Bureau 2019)
- Geographic Mobility (percent of People Who Moved in During the Past Year) (United States Census Bureau 2019)
- Average value of Single, Detached Housing Units (United States Census Bureau 2019)

CATEGORICAL

- Majority White or Not Majority White Population (United States Census Bureau 2020)
- Urban Rural Classification (National Center for Health Statistics 2013)

## data table
The dataset includes 1478 census tracks. Note - the housing value is only available for 337 out of your 1478 census tracts, so we have filtered the data to only include those, which substantially reduces the sample size.

```{r load library, message=FALSE, warning=FALSE, echo = FALSE, include = FALSE}
library(tidyverse)
library(tidycensus)
library(readxl)
library(knitr)
library(sf)
library(gridExtra)
library(DT)
```


````{r run variables, message = FALSE, include = FALSE, echo = FALSE, warning = FALSE}
acs_vars <- c(pop = "B01003_001",
          HUs = "B25001_001",
          oneunit = "B25024_002",
          one_house_value = "B25080_002",
          tot_value = "B25080_001",
          w_pop = 'B02001_002',
          moved_same_county = "B07001_033",
          moved_diff_state = "B07001_065",
          moved_same_state = "B07001_049",
          moved_abroad = "B07001_081")

census <- get_acs(geography = "tract", 
                        year = 2019,
                        state = "MA",
                        variables = acs_vars,
                        output = "wide") %>%
  mutate(pct_sf_home = oneunitE / HUsE) %>%
  mutate(pct_moved = (moved_same_countyE +
           moved_diff_stateE +
           moved_same_stateE +
           moved_abroadE) / HUsE) %>%
  mutate(avg_sf_home_val = one_house_valueE / oneunitE) %>%
  mutate(pct_w = w_popE / popE) %>%
  mutate(maj_white = ifelse(pct_w > 0.5, TRUE, FALSE)) %>%
  select(GEOID, pct_sf_home, pct_moved, avg_sf_home_val, maj_white) %>%
  filter(!is.na(avg_sf_home_val)) %>%
  mutate(GEOID_county = substr(GEOID, 1, 5))
```


```{r calculate urban type, message = FALSE, warning = FALSE, echo = FALSE, include = FALSE}
#give the GEOID column you're creating the same name as the county GEOID in your census dataset.
urban_type <- read_xlsx(path = "data/NCHSURCodes2013.xlsx", 
                      sheet = "NCHSURCodes2013") %>%
  mutate(GEOID_county = 
           case_when(str_length(as.character(`FIPS code`)) == 5 ~ 
                            as.character(`FIPS code`),
                          str_length(as.character(`FIPS code`)) == 4 ~
                            paste("0", `FIPS code`, sep=""),
                          TRUE ~ "unknown")) %>%
  mutate(type = case_when(`2013 code` == 1 ~ "Large central metro",
                          `2013 code` == 2 ~ "Large fringe metro",
                          `2013 code` == 3 ~ "Medium metro",
                          `2013 code` == 4 ~ "Small metro",
                          `2013 code` == 5 ~ "Micropolitan",
                          `2013 code` == 6 ~ "Non-core",
                          TRUE ~ "unknown")) %>%
  select(GEOID_county, type)
```

```{r create data table with kable, message = FALSE, warning = FALSE,echo = FALSE}
#join the types to the census tracts
#view the first few rows of this combined data set.

data <- left_join(census,urban_type) 

kable(head(data))

```
This dataset includes 337 variables.

# descriptive statistics

## continuous variables
The continuous variables we use are: 1) percentage of single, detached housing units, 2) geographic mobility, 3) average value of single, detached housing units.

```{r t_tests, message = FALSE, echo = FALSE, warning = FALSE}
pct_moved_t_test <- t.test(data$pct_moved)
pct_sf_home_t_test <- t.test(data$pct_sf_home)
avg_sf_home_val_t_test <- t.test(data$avg_sf_home_val)

#Quartile
pct_moved_quartiles <- quantile(data$pct_moved, na.rm = TRUE)
#Standard Deviation
pct_moved_st_dev <- sd(data$pct_moved, na.rm = TRUE)
#Histogram
pct_moved_hist <- ggplot(data) +
  geom_histogram(aes(x = pct_moved))
#Quartile
pct_sf_home_quartiles <- quantile(data$pct_sf_home, na.rm = TRUE)
#Standard Deviation
pct_sf_home_st_dev <- sd(data$pct_moved, na.rm = TRUE)
#Histogram
pct_sf_home_hist <- ggplot(data) +
  geom_histogram(aes(x = pct_sf_home))
#Quartile
avg_sf_home_val_quartiles <- quantile(data$avg_sf_home_val, na.rm = TRUE)
#Standard Deviation
avg_sf_home_val_st_dev <- sd(data$avg_sf_home_val, na.rm = TRUE)
#Histogram
avg_sf_home_val_hist <- ggplot(data) +
  geom_histogram(aes(x = avg_sf_home_val))
```

```{r summary table continuous}
cont_summary <- tibble(
  variable = c("Percent of Single, Detached Housing Units (%)", 
               "Percent of Residents Who Arrived in Past Year (%)", 
               "Average Home Value for Single, Detached Housing Units"),
  
  `sample mean` = c(pct_moved_t_test$estimate,
                    pct_sf_home_t_test$estimate*100,
                    avg_sf_home_val_t_test$estimate),
  
  `population mean (95% confidence) - low` = 
    c(pct_moved_t_test$conf.int[1],
      pct_sf_home_t_test$conf.int[1]*100,
       avg_sf_home_val_t_test$conf.int[1]),
  
  `population mean (95% confidence) - high` =
    c(pct_moved_t_test$conf.int[2],
      pct_sf_home_t_test$conf.int[2]*100,
      avg_sf_home_val_t_test$conf.int[2]),
  
  median = c(pct_moved_quartiles[3],
             pct_sf_home_quartiles[3]*100,
             avg_sf_home_val_quartiles[3]),
  
  `interquartile range` = c((pct_moved_quartiles[4] - pct_moved_quartiles[2]),
                            (pct_sf_home_quartiles[4] - pct_sf_home_quartiles[2])*100,
                            avg_sf_home_val_quartiles[4] - avg_sf_home_val_quartiles[2]),
  
  `standard deviation` = c(pct_moved_st_dev,
                          pct_sf_home_st_dev*100,
                          avg_sf_home_val_st_dev))
df <- cont_summary 

library(formattable)
cont_summary2 <- currency(df$3, digits = 0L)


kable(cont_summary, digits = 2)
```

```{r}
pct_moved_hist <- ggplot(data) +
  geom_histogram(aes(x = pct_moved))


pct_sf_home_hist <- ggplot(data) +
  geom_histogram(aes(x = pct_sf_home))


avg_sf_home_val_hist <- ggplot(data) +
  geom_histogram(aes(x = avg_sf_home_val), bins = 30)


pretty_pct_moved_hist <- pct_moved_hist +
  theme_bw() +
  scale_x_continuous(name = "Percent of people who arrived\nin past year") +
  scale_y_continuous(name = "Number of census tracts") +
  theme(axis.text.x = element_text(angle = 90))

pretty_avg_sf_home_val_hist <- avg_sf_home_val_hist +
  theme_bw() + 
  scale_x_continuous(name = "Average home value for\nsingle, detached housing units") +
  scale_y_continuous(name = "Number of census tracts") +
  theme(axis.text.x = element_text(angle = 90))

pretty_pct_sf_home_hist = pct_sf_home_hist +
  theme_bw() +
  scale_x_continuous(name = "Percent of single, detached\nhousing units") +
  scale_y_continuous(name = "Number of census tracts") +
  theme(axis.text.x = element_text(angle = 90)) 

grid.arrange(pretty_pct_moved_hist, pretty_avg_sf_home_val_hist, pretty_pct_sf_home_hist,
             ncol = 3)
```


## categorical Variables
The categorical variables we use are: 1) Majority White or Non-Majority White Classification of Census Tract and 2) Urban-Rural Classification

```{r t_tersts, message = FALSE, echo = FALSE, warning = FALSE, include = FALSE}
pct_wm <- t.test(data$maj_white == "TRUE")
pct_wnm <-  t.test(data$maj_white == "FALSE")
pct_large_metro <- t.test(data$type == "Large central metro")
pct_fringe_metro <- t.test(data$type == "Large fringe metro")
pct_med_metro <- t.test(data$type == "Medium metro")
pct_small_metro <- t.test(data$type == "Small metro")
pct_micro <- t.test(data$type == "Micropolitan")
pct_rural <- t.test(data$type == "Non-core")
```

### census tracts with a majority white population
insert description

```{r majority white summary, message = FALSE, echo = FALSE, warning = FALSE}
categorical_summary_white <- tibble(`Racial Composition of Census Tract` = 
                              c("Majority White",
                                "Majority Non-White"),
                            `Sample Proportion` = 
                              c(pct_wm$estimate * 100,
                                pct_wnm$estimate * 100),
                            `95-percent confidence - low` = 
                              c(pct_wm$conf.int[1] * 100,
                                pct_wnm$conf.int[1] * 100),
                            `95-percent confidence - high` = 
                              c(pct_wm$conf.int[2] * 100,
                                pct_wnm$conf.int[2] * 100))

kable(categorical_summary_white, digits = 0)
```

```{r majority white plot, message = FALSE, echo = FALSE, warning = FALSE }
ggplot(categorical_summary_white) +
  geom_bar(aes(x = `Racial Composition of Census Tract`, 
               y = `Sample Proportion`),
           stat = "identity") +
  geom_errorbar(aes(x = `Racial Composition of Census Tract`, 
               y = `Sample Proportion`,
               ymin = `95-percent confidence - low`,
               ymax = `95-percent confidence - high`),
           stat = "identity") +
  scale_y_continuous(name = "Percent of Census Tracts",
                     breaks = c(0, 20, 40, 60, 80),
                     labels = c("0", "20%", "40%", "60%", "80%")) +
  theme_bw()
```

### urban-rural classifications
insert text here

```{r urban rural summary, message = FALSE, echo = FALSE, warning = FALSE}
categorical_summary_type <- tibble(`County type` = 
                              c("Larch central metro",
                                "Large fringe metro",
                                "Medium metro",
                                "Small metro",
                                "Micropolitan",
                                "Non-core"),
                            `Sample proportion` = 
                              c(pct_large_metro$estimate * 100,
                                pct_fringe_metro$estimate * 100,
                                pct_med_metro$estimate * 100,
                                pct_small_metro$estimate * 100,
                                pct_micro$estimate * 100,
                                pct_rural$estimate * 100),
                            `95-percent confidence - low` = 
                              c(pct_large_metro$conf.int[1] * 100,
                                pct_fringe_metro$conf.int[1] * 100,
                                pct_med_metro$conf.int[1] * 100,
                                pct_small_metro$conf.int[1] * 100,
                                pct_micro$conf.int[1] * 100,
                                pct_rural$conf.int[1] * 100),
                            `95-percent confidence - high` = 
                              c(pct_large_metro$conf.int[2] * 100,
                                pct_fringe_metro$conf.int[2] * 100,
                                pct_med_metro$conf.int[2] * 100,
                                pct_small_metro$conf.int[2] * 100,
                                pct_micro$conf.int[2] * 100,
                                pct_rural$conf.int[2] * 100))

kable(categorical_summary_type, digits = 0)
```

```{r urban rural plot, message = FALSE, echo = FALSE, warning = FALSE}
ggplot(categorical_summary_type) +
  geom_bar(aes(x = `County type`, 
               y = `Sample proportion`),
           stat = "identity") +
  geom_errorbar(aes(x = `County type`, 
               y = `Sample proportion`,
               ymin = `95-percent confidence - low`,
               ymax = `95-percent confidence - high`),
           stat = "identity") +
  scale_y_continuous(name = "Percent of counties",
                     breaks = c(0, 10, 20, 30, 40, 50, 60, 70),
                     labels = c("0", "10%", "20%", "30%", "40%", "50%", "60%", "70%")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) 
```


