---
title: "ses5215-assignment01-Megan"
author: "Mike-Lid"
date: "10/21/2021"
output: html_document
---

```{r Loading libraries}

library(tidyverse)
library(tidycensus)
library(readxl)
library(knitr)
```

```{r}
acs_vars <- load_variables(2019, "acs5")
```

```{r}
tot_moved <- get_acs(geography = "tract", state = "Massachusetts",  
                         variables = c(same_county = "B07001_033", diff_state = "B07001_065", 
                                       same_state = "B07001_049", abroad = "B07001_081"), output = 
                          "wide", quiet = TRUE)
```

```{r}

tot_moved$tot_moved <- moved$same_countyE + moved$same_stateE + moved$diff_stateE + moved$abroadE

  
```
### Assignment 2
### Running T Test - Confidence Interval
```{r}
pct_moved_t_test <- t.test(data$pct_moved)
pct_sf_home_t_test <- t.test(data$pct_sf_home)
avg_sf_home_val_t_test <- t.test(data$avg_sf_home_val)
```

## Quartiles
```{r}
pct_moved_quartiles <- quantile(data$pct_moved, na.rm = TRUE)
pct_sf_home_quartiles <- quantile(data$pct_sf_home, na.rm = TRUE)
avg_sf_home_val_quartiles <- quantile(data$avg_sf_home_val, na.rm = TRUE)

```

## Standard Deviation
```{r}
pct_moved_st_dev <- sd(data$pct_moved, na.rm = TRUE)
pct_sf_home_st_dev <- sd(data$pct_moved, na.rm = TRUE)
avg_sf_home_val_st_dev <- sd(data$avg_sf_home_val, na.rm = TRUE)


```

## Histograms
```{r}
pct_moved_hist <- ggplot(data) +
  geom_histogram(aes(x = pct_moved))

pct_moved_hist

pct_sf_home_hist <- ggplot(data) +
  geom_histogram(aes(x = pct_sf_home))

pct_sf_home_hist

avg_sf_home_val_hist <- ggplot(data) +
  geom_histogram(aes(x = avg_sf_home_val), bins = 30)

avg_sf_home_val_hist
```

### Correlation - Home Value Correlated with Geographic Mobility
```{r}
cor.test(~ avg_sf_home_val + pct_moved, data = data)
```
Running the correlation shows that there is a positive correlation between the two variables-- the percent of people who moved to a census tract and the average value of single family homes. With 95 percent confidence, the value of single-family homes is correlated with the percent of people moving to a census tract.

### Correlation - Home Value Correlated with Geographic Mobility
```{r}
cor.test(~ avg_sf_home_val + pct_sf_home, data = data)

```

