---
title: "ses5215-assignment01-Megan"
author: "Mike-Lid"
date: "10/21/2021"
output: html_document
---

```{r Loading libraries}

library(tidyverse)
library(tidycensus)
library(readxl)
library(knitr)
```

```{r}
acs_vars <- load_variables(2019, "acs5")
```

```{r}
tot_moved <- get_acs(geography = "tract", state = "Massachusetts",  
                         variables = c(same_county = "B07001_033", diff_state = "B07001_065", 
                                       same_state = "B07001_049", abroad = "B07001_081"), output = 
                          "wide", quiet = TRUE)
```

```{r}

tot_moved$tot_moved <- moved$same_countyE + moved$same_stateE + moved$diff_stateE + moved$abroadE

  
```
### Assignment 2
### Running T Test - Confidence Interval
```{r}
pct_moved_t_test <- t.test(data$pct_moved)
pct_sf_home_t_test <- t.test(data$pct_sf_home)
avg_sf_home_val_t_test <- t.test(data$avg_sf_home_val)
```

## Quartiles
```{r}
pct_moved_quartiles <- quantile(data$pct_moved, na.rm = TRUE)
pct_sf_home_quartiles <- quantile(data$pct_sf_home, na.rm = TRUE)
avg_sf_home_val_quartiles <- quantile(data$avg_sf_home_val, na.rm = TRUE)

```

## Standard Deviation
```{r}
pct_moved_st_dev <- sd(data$pct_moved, na.rm = TRUE)
pct_sf_home_st_dev <- sd(data$pct_moved, na.rm = TRUE)
avg_sf_home_val_st_dev <- sd(data$avg_sf_home_val, na.rm = TRUE)


```

## Histograms
```{r}
pct_moved_hist <- ggplot(data) +
  geom_histogram(aes(x = pct_moved))

pct_moved_hist

pct_sf_home_hist <- ggplot(data) +
  geom_histogram(aes(x = pct_sf_home))

pct_sf_home_hist

avg_sf_home_val_hist <- ggplot(data) +
  geom_histogram(aes(x = avg_sf_home_val), bins = 30)

avg_sf_home_val_hist
```

### Correlation - Home Value Correlated with Geographic Mobility
```{r}
cor.test(~ avg_sf_home_val + pct_moved, data = data)
```
Running the correlation shows that there is a positive correlation between the two variables-- the percent of people who moved to a census tract and the average value of single family homes. With 95 percent confidence, the value of single-family homes is correlated with the percent of people moving to a census tract.

### Correlation - Home Value Correlated with Geographic Mobility
```{r}
cor.test(~ avg_sf_home_val + pct_sf_home, data = data)

```




```{r}
coeff_names_logged <- c("Constant" = "(Intercept)",
                 
                 "Percentage of SDDU" = "pct_sf_home",
                 "Percentage of SDDU (Logged)" = "log2_pct_sf_home",
                 "Percentage of Moved (5% Increments)" = "pct_moved_5",
                 "Percentage of Moved" = "pct_moved",
                 "Percentage of Moved (Logged)" = "log2_pct_moved",
                 "Majority White Classification" = "maj_whiteTRUE",
                 "Large fringe metro" = "typeLarge fringe metro",
                 "Medium metro" = "typeMedium metro",
                 "Small metro" = "typeSmall metro",
                 "Micropolitan" = "typeMicropolitan")
                   
                   
model_log_dependent <- lm(log2_avg_sf_home_val ~ log2_pct_sf_home + log2_pct_moved + maj_white + type, data_log)
model_log <- lm(avg_sf_home_val ~ log2_pct_sf_home + log2_pct_moved + maj_white + type, data_log)

export_summs(full_model, centered_model, centered_model_scaled, model_log_dependent, model_log,
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Initial\n(price in dollars)", 
                             "Centered\n(price in dollars)", 
                             "Centered+Scaled\n(price in thousand of dollars)",
                             "Logged dependent variable\n(price in dollars)",
                             "Logged only independent variables\n(price in dollars)"),
             coefs = coeff_names_logged)

```





## Assignment 5 Intro Language

# **transformations**
We applied simple *linear* transformations to our data set in the below analyses, where we multiplied our percent values by a constant value  (100) so that the decimal format could be communicated in terms of percentage (%). This made our values easier to interpret since it relied on whole numbers and not a string of decimals. We also experimented with *nonlinear* transformations and discovered that using a logarithmic transformation on our dependent variable helped improve the overall model fit for our data.

## **mean centering**

We mean centered the model to make our intercept easier to interpret. Our initial intercept indicated the predicted increase in the values of single-family detached units (SDDUs) for census tracts with 0% SDDUs and in which 0% of the population had moved in within the last year (geographic mobility). This would be a highly unlikely--or at least very unusual--census tract. By means centering, our intercept shows that the predicted increase in unit value for SDDUs is approximately $303,153 for census tracts with an average percentage of SDDUs and average geographic mobility. Interestingly, this was not dramatically different than the original, uncentered intercept (307,328), perhaps indicating that the "average" census tract with respect to our variables does have a fairly small percentage of SDDUs and geographic mobility. 

```{r, include = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
lm(avg_sf_home_val ~ pct_sf_home + pct_moved + maj_white + type, data) 
```

```{r, include = FALSE, warning = FALSE, message = FALSE, echo = FALSE}
centered_data <- data %>%
  mutate(pct_sf_home = pct_sf_home - mean(pct_sf_home, na.rm=TRUE),
         pct_moved = pct_moved - mean(pct_moved, na.rm=TRUE))


centered_model <- lm(avg_sf_home_val ~ pct_sf_home + pct_moved + maj_white + type, centered_data)

export_summs(full_model, centered_model, 
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Initial", "Centered"))
```


```{r}
coeff_names <- c("Constant" = "(Intercept)",
                 "Percentage of Single, Detached Dwelling  Units" = "pct_sf_home",
                 "Percentage of Population Moved" = "pct_moved",
                 "Majority White Classification" = "maj_whiteTRUE",
                 "Large fringe metro" = "typeLarge fringe metro",
                 "Medium metro" = "typeMedium metro",
                 "Small metro" = "typeSmall metro",
                 "Micropolitan" = "typeMicropolitan")

export_summs(full_model, centered_model, 
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Initial", "Centered"),
             coefs = coeff_names)
```


## **scaling**

We scaled our data by dividing the average price of SDDUs in a census tract by 1,000 because we are interested in grouping unit prices by thousands of dollars. Differences in prices by single dollars is less meaningful given that we are working with values in the hundreds of thousands. We chose not to scale the percentage of SDDUs in a census tract because we are interested to see the effect of these by each percentage. We did, however, scale the percentage of residents moved by dividing by 5, so that we could categorize the percentages into 5% increments.

```{r}
centered_data_scaled <- centered_data %>%
  mutate(avg_sf_home_val_1k = avg_sf_home_val / 1000,
         pct_moved_5 = pct_moved / 5)

#scaling to change from calculating how these variables effect the dollars of Single, Detached Dwelling Units (SDDU) to how these variables effects the price in terms of 1000's of dollars.

coeff_names_scaled <- c("Constant" = "(Intercept)",
                 "Percentage of Single, Detached Dwelling  Units" = "pct_sf_home",
                 "Percentage of Population Moved (5% Increments)" = "pct_moved_5",
                 "Majority White Classification" = "maj_whiteTRUE",
                 "Large fringe metro" = "typeLarge fringe metro",
                 "Medium metro" = "typeMedium metro",
                 "Small metro" = "typeSmall metro",
                 "Micropolitan" = "typeMicropolitan")
                   
                   
centered_model_scaled <- lm(avg_sf_home_val_1k ~ pct_sf_home + pct_moved_5 + maj_white + type, centered_data_scaled)

export_summs(full_model, centered_model, centered_model_scaled,
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Initial\n(price in dollars)", 
                             "Centered\n(price in dollars)", 
                             "Centered\n(price in thousand of dollars)"),
             coefs = coeff_names_scaled)


```

## **nonlinear transformations**
We were initially unsure whether any of our variables were likely to have non-linear relationships, so we did some experimentation with logarithmic transformations to see whether they improved our model fit. We suspected that our dependent variable--SDDU prices--could potentially have a non-linear relationship. In fact, doing a base 2 log transformation of our dependent variable and the percentage of population moved to a census tract did improve the fit of our model (from 0.19 to 0.25). This transformation complicates our interpretation somewhat, because we are now predicting based on the change in the *log* of the SDDU price. Nonetheless, the model is a better fit for the data, and the regression indicates a statistically significant, positive effect on SDDU prices ceteris paribus.

Logging the percentage of SDDUs in a given census tract did not effect the R2 value, but did change some of the level of significance for variables. Therefore, because the model fit is equally good, we selected the regression model that has each of our continuous variables transformed.    

```{r, message = FALSE, warning = FALSE, echo = FALSE}
#logging the percentage of sf homes did not seem to affect it
data_log <- data %>%
  mutate(log2_avg_sf_home_val = log(avg_sf_home_val + mean(data$avg_sf_home_val, na.rm = TRUE), base = 2),
         log2_pct_moved = log(pct_moved + mean(data$pct_moved, na.rm = TRUE), base = 2),
         log2_pct_sf_home = log(pct_sf_home + mean(data$pct_sf_home, na.rm = TRUE), base = 2))
```



## Assignment 5 Summary
To arrive at this alternative version of our regression, we experimented with both linear and nonlinear transformations.  

First, we scaled our data by dividing the average price of SDDUs in a census tract by 1,000 because we are interested in grouping unit prices by thousands of dollars. Differences in prices by single dollars is less meaningful given that we are working with values in the hundreds of thousands. We chose not to scale the percentage of SDDUs in a census tract because we are interested to see the effect of these by each percentage. We did, however, scale the percentage of residents moved by dividing by 5, so that we could categorize the percentages into 5% increments.

We were initially unsure whether any of our variables were likely to have non-linear relationships, so we did some experimentation with logarithmic transformations to see whether they improved our model fit. We suspected that our dependent variable--SDDU prices--could potentially have a non-linear relationship. In fact, doing a base 2 log transformation of our dependent variable and the percentage of population moved to a census tract did improve the fit of our model (from 0.19 to 0.25). This transformation complicates our interpretation somewhat, because we are now predicting based on the change in the *log* of the SDDU price. Nonetheless, the model is a better fit for the data, and the regression indicates a statistically significant, positive effect on SDDU prices ceteris paribus.

Logging the percentage of SDDUs in a given census tract did not effect the R2 value, but did change some of the level of significance for variables. Therefore, because the model fit is equally good, we selected the regression model that has each of our continuous variables transformed. 

## Scaling
```{r}
data_scaled <- data %>%
  mutate(avg_sf_home_val_1k = avg_sf_home_val / 1000,
         pct_moved_5 = pct_moved / 5)

#scaling to change from calculating how these variables effect the dollars of Single, Detached Dwelling Units (SDDU) to how these variables effects the price in terms of 1000's of dollars.

coeff_names_scaled <- c("Constant" = "(Intercept)",
                 "Percentage of Single, Detached Dwelling  Units" = "pct_sf_home",
                 "Percentage of Population Moved (5% Increments)" = "pct_moved_5",
                 "Majority White Classification" = "maj_whiteTRUE",
                 "Large fringe metro" = "typeLarge fringe metro",
                 "Medium metro" = "typeMedium metro",
                 "Small metro" = "typeSmall metro",
                 "Micropolitan" = "typeMicropolitan")
                   
                   
model_scaled <- lm(avg_sf_home_val_1k ~ pct_sf_home + pct_moved_5 + maj_white + type, data_scaled)

export_summs(full_model, model_scaled,
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Initial\n(price in dollars)", 
                             "Scaled\n(price in thousand of dollars)"),
             coefs = coeff_names_scaled)
```

## Logarithimic transformations
```{r, message = FALSE, warning = FALSE, echo = FALSE}
#logging the percentage of sf homes did not seem to affect it
data_scaled_log <- data_scaled %>%
  mutate(log2_avg_sf_home_val = log(avg_sf_home_val, base = 2),
         log2_pct_moved = log(pct_moved, base = 2),
         log2_pct_sf_home = log(pct_sf_home, base = 2))
```


```{r}
coeff_names_logged <- c("Constant" = "(Intercept)",
                 
                 "Percentage of SDDU" = "pct_sf_home",
                 "Percentage of SDDU (Logged)" = "log2_pct_sf_home",
                 "Percentage of Moved (5% Increments)" = "pct_moved_5",
                 "Percentage of Moved" = "pct_moved",
                 "Percentage of Moved (Logged)" = "log2_pct_moved",
                 "Majority White Classification" = "maj_whiteTRUE",
                 "Large fringe metro" = "typeLarge fringe metro",
                 "Medium metro" = "typeMedium metro",
                 "Small metro" = "typeSmall metro",
                 "Micropolitan" = "typeMicropolitan")
                   
                   
model_log_dependent <- lm(log2_avg_sf_home_val ~ log2_pct_sf_home + log2_pct_moved + maj_white + type, data_log)
model_log <- lm(avg_sf_home_val ~ log2_pct_sf_home + log2_pct_moved + maj_white + type, data_log)

export_summs(full_model, centered_model, centered_model_scaled, model_log_dependent, model_log,
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Initial\n(price in dollars)", 
                             "Centered\n(price in dollars)", 
                             "Centered+Scaled\n(price in thousand of dollars)",
                             "Logged dependent variable\n(price in dollars)",
                             "Logged only independent variables\n(price in dollars)"),
             coefs = coeff_names_logged)

```

### Interactions
```{r}
coeff_names_interaction <- c("Constant" = "(Intercept)",
                               "Percentage of SDDU" = "pct_sf_home",
                 "Percentage of SDDU (Logged)" = "log2_pct_sf_home",
                 "Percentage of Moved (5% Increments)" = "pct_moved_5",
                 "Percentage of Moved" = "pct_moved",
                 "Percentage of Moved (Logged)" = "log2_pct_moved",
                 "Majority White Classification" = "maj_whiteTRUE",
                 "Large fringe metro" = "typeLarge fringe metro",
                 "Medium metro" = "typeMedium metro",
                 "Small metro" = "typeSmall metro",
                 "Micropolitan" = "typeMicropolitan")
```

```{r}
interaction_model_log_dependent <- lm(log2_avg_sf_home_val ~ log2_pct_sf_home + log2_pct_moved + maj_white + type + log2_pct_sf_home*maj_white + log2_pct_sf_home*type + log2_pct_moved*maj_white + log2_pct_moved*type, data_log)

export_summs(centered_model_scaled_log, interaction_model,
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Model 4",
                             "Model 5"))
```


